Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-19 WholePlotTrajectories.Rmd
citatoin check

keywords: [Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience]
journalinfo: Hal xxx
archive: DOI xxxx
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-ga-ni-za-tion}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::gitbook: default
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
---

@@

```{r Options, echo=FALSE, message=FALSE, warning=F}
knitr::opts_chunk$set(cache=T, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
Library("kableExtra")

@@

The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1989;@Blanc2009]. 
The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found) measured with the BIOMASS R package [@Biomass2018].

@@

This variability of protocols raised methodological issues as vernacular names usually correspond to different botanical species. 
It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
Vernacular names were replaced through multinomial trials $M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)$ based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$. 
See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.

@@

## Communities composition

@@

Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped shaped trajectories (\@ref(fig:RedFun_rest)).
The maximum redundancy loss was positively correlated with the disturbance intensity (XX spearman to be measured) and the initial value had not recovered for any disturbed communities.

```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial communities functional space over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
load("DB/Redundancy_restricted")

>>>>

keywords: [Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience]
journalinfo: Hal xxx
archive: DOI xxxx
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
  bookdown::word_document2: default
---

@@

```{r Options, echo=FALSE, message=FALSE, warning=F}
knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
Library("kableExtra")

@@

The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1990;@Blanc2009]. 
The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found) measured with the BIOMASS R package [@Biomass2018].

@@

This variability of protocols raised methodological issues as vernacular names usually correspond to different botanical species. 
It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$:

\begin{align}
M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
\end{align}

See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.

@@

## Communities Composition

@@

Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped shaped trajectories (\@ref(fig:RedFun_rest)).
The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.50$) and the initial value had not recovered for any disturbed communities.

```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
load("DB/Redundancy_restricted")

### 2018-06-19 WholePlotTrajectories.tex
citatoin check

% User-adder preamble
\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
\usepackage{tabu} \usepackage{caption}
\captionsetup{justification = justified}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings}

@@

The experiment is a network of twelve 6.25ha plots that underwent a
gradient of three logging, thinning and fuelwood cutting treatments
(Table \ref{tab:Tab1}) attributed according to a randomized plot design
with three replicate blocks of four plots. The disturbance corresponds
to averages of 10 trees removed per hectare with a diameter at 1.3 m
height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
(T3). Treatments T2 and T3 besides included the thinning of trees by
poison girdling \citep{Schmitt1989, Blanc2009}. The disturbance
intensity was measured as the percentage of aboveground biomass (\%AGB)
lost between the first inventory in 1984 and five years after
disturbance (ref to be found) measured with the BIOMASS R package
\citep{Biomass2018}.

@@

This variability of protocols raised methodological issues as vernacular
names usually correspond to different botanical species. It resulted in
significant taxonomic uncertainties that had to be propagated to
composition and diversity metrics. The uncertainty propagation was done
through a Bayesian framework reconstituting complete inventories at
genus level from real incomplete ones on the basis of
vernacular/botanical names association. Vernacular names were replaced
through multinomial trials
\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
based on the association probability
\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) observed across all
inventories between each vernacular name \emph{v} and the species
\(\big[s_1, s_2, …, s_N\big]\). See appendix 1 and
\citet{Aubry-Kientz2013} for the detailed methodology.

@@

\subsection{Communities composition}\label{communities-composition}

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life trait (Specific maximum height at adult stage, \emph{Hmax}). Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
\end{figure*}

@@

Communities functional redundancy was measured as the sum within
communities the species weighted functional overlap based on the 7 leaf,
stem, and maximum height traits (see appendix I for PCA details).
Communities functional redundancy remained stable in control plots but
after disturbance the redundancy trajectories were quite variable (See
appendix I) and apparently independently of the initial disturbance.
Globally after most intense disturbance (plots T2 and T3) communities
redundancy decreased at first place before increasing to edge, recover
or exceed the initial value.

Considering the functional redundancy restricted to the functional space
of the initial inventory, all disturbed plots followed similar
decreasing humped shaped trajectories (@ref(fig:RedFun\_rest)). The
maximum redundancy loss was positively correlated with the disturbance
intensity (XX spearman to be measured) and the initial value had not
recovered for any disturbed communities.

@@

\caption{Trajectories of the functional redundancy within the initial cpùùunities functional space over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}(\#fig:RedFun_rest)
\end{figure}

@@

The examination of taxonomic and functional trajectories suggested that
communities functional structure and dominant species pool relied upon
deterministic processes following the IDH while the pool of rare species
relied upon more stochastic processes.

The disturbance intensity did not predict the taxonomic richness, as
already observed in the Paracou station \citep{Baraloto2012a} and in
Bornean tropical forests \citep{Cannon1998}, nor did it strongly
correlate with the taxonomic evenness (\emph{i.e.} Simpson diversity)
although the it was enhanced after disturbance.

Contrastingly, the disturbance intensity predicted the increase in
functional diversity along the 30 years monitored after disturbance. The
increase in functional diversity paralleled the important specific
turnover and significant functional shifts towards resource-acquisitive
strategies (sharp increase in the SLA, leaf thickness and bark thickness
and decrease in wood density, leaf toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
pre-disturbance survivor trees proved to mirror initial communities
\citep{Herault2018} the functional trajectories and taxonomic turnover
rather relied on the pool of newly recruited trees driven by the
enhanced growth and survival of previously infrequent species and
functional types. Disturbance entailed a reorganization of the typical
high dominance structure of hyperdiverse mature forests benefiting to
pioneers and light demanding species: the changes in abiotic environment
and competitive pressure favored pioneers which outcompete other species
in non limiting resources but were excluded in mature forests by
long-lived, more resistant and shade tolerant species. Disturbance
trajectories therefore relied on the environmental niches made available
and filled by species from a restricted functional range that became
dominant and structured the functionalcharacteristics of the community
\citep{Grime1998}. The intermediate disturbance hypothesis predicted
communities functional diversity and pool of dominant species
\citep{Molino2001}, but rare species relied on different processes and
disprove the IDh for the whole taxonomic community
\citep{Hubbell2001, Chave2004}.

@@

The functional and taxonomic decoupling and the disprove of the IDH for
communities taxonomic structure were explained by persisting rearranging
of communities functional redundancy that implied an incomplete
resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.

>>>>

% User-adder preamble
\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
\usepackage{tabu} \usepackage{caption}
\captionsetup{justification = justified}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion}

@@

The experiment is a network of twelve 6.25ha plots that underwent a
gradient of three logging, thinning and fuelwood cutting treatments
(Table \ref{tab:Tab1}) attributed according to a randomized plot design
with three replicate blocks of four plots. The disturbance corresponds
to averages of 10 trees removed per hectare with a diameter at 1.3 m
height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
(T3). Treatments T2 and T3 besides included the thinning of trees by
poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
intensity was measured as the percentage of aboveground biomass (\%AGB)
lost between the first inventory in 1984 and five years after
disturbance (ref to be found) measured with the BIOMASS R package
\citep{Biomass2018}.

@@

This variability of protocols raised methodological issues as vernacular
names usually correspond to different botanical species. It resulted in
significant taxonomic uncertainties that had to be propagated to
composition and diversity metrics. The uncertainty propagation was done
through a Bayesian framework reconstituting complete inventories at
genus level from real incomplete ones on the basis of
vernacular/botanical names association. Vernacular names were replaced
through multinomial trials based on the association probability
\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) observed across all
inventories between each vernacular name \emph{v} and the species
\(\big[s_1, s_2, …, s_N\big]\):

\begin{align}
M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
\end{align}

See appendix 1 and \citet{Aubry-Kientz2013} for the detailed
methodology.

@@

\subsection{Communities Composition}\label{communities-composition}

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
\end{figure*}

@@

Communities functional redundancy was measured as the sum within
communities of species functional overlap weighted by their abundance.
The functional redundancy was measured in the functional space defined
by 7 leaf, stem, and maximum height traits (see appendix I for PCA
details). Communities functional redundancy remained stable in control
plots but after disturbance the redundancy trajectories were quite
variable (See appendix I) and apparently independently of the initial
disturbance. Globally after most intense disturbance (plots T2 and T3)
communities redundancy decreased at first place before increasing to
edge, recover or exceed the initial value.

Considering the functional redundancy restricted to the functional space
of the initial inventory, all disturbed plots followed similar
decreasing humped shaped trajectories (@ref(fig:RedFun\_rest)). The
maximum redundancy loss was positively correlated with the disturbance
intensity (\(\rho_{spearman}=0.50\)) and the initial value had not
recovered for any disturbed communities.

@@

\caption{Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}(\#fig:RedFun_rest)
\end{figure}

@@

The taxonomic and functional trajectories validated the IDH as
determinant of communities functional structure and of the dominant
species pool, while contrastingly the pool of rare species relied upon
more stochastic processes.

The disturbance intensity poorly predicted the taxonomic richness, as
already observed in the Paracou station \citep{Baraloto2012a} and in
Bornean tropical forests \citep{Cannon1998}, was only weakly correlated
with the increase in taxonomic evenness (\emph{i.e.} Simpson diversity).

Contrastingly, the disturbance intensity consistently predicted the
increase in functional diversity after disturbance along the 30 years.
The functional diversity increase paralleled the significant specific
turnover and functional shifts towards resource-acquisitive strategies
(sharp increase in the SLA, leaf thickness and bark thickness and
decrease in wood density, leaf toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
pre-disturbance survivor trees proved mirroring the initial communities
\citep{Herault2018} the functional trajectories and taxonomic turnover
rather relied upon the pool of newly recruited trees driven by the
enhanced growth and survival of previously infrequent species and
functional types. Disturbance entailed a reorganization, benefiting to
pioneers and light demanding species, of the typical high dominance
structure of hyperdiverse mature forests. The changes in abiotic
environment and competitive pressure favored pioneers which outcompeted
other species in non limiting light resources but were excluded in
mature forests by long-lived, more resistant and shade tolerant species.
Disturbance trajectories therefore relied upon the environmental niches
made available and filled by species from a restricted functional range
that became dominant and structured the functional characteristics of
the community \citep{Grime1998}. The intermediate disturbance hypothesis
was suitabe to predict communities functional diversity and pool of
dominant species \citep{Molino2001}, but it poorly represented the whole
taxonomic structure that was hampered by the slow recovery of rare
species \citep{Hubbell2001, Chave2004}.

@@

The functional and taxonomic decoupling and the rebuttal of the IDH for
communities taxonomic structure were explained by a reorganization of
the functional redundancy within the functional traits space that
implied the incomplete resilience of communities
\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
